\section{Related Work}

Previous work has been done on measuring the power and performance characteristics of 3G RRC state machines~\cite{3g_rrc} as well as 4G LTE networks~\cite{4g_rrc, drx_analysis} in controlled environments, as well as specific features of those networks such as DRX~\cite{drx}.  Work has also been done on exploring the implications on application performance of RRC state and inadequacies in how mobile applications deal with RRC state~\cite{aro}.  

We expand upon this work by implementing an RRC inference method that can be integrated into a measurement tool intended for non-experts. It can automatically infer the RRC state machine of any network type, allowing anomalous or unexpected behavior to be measured, unlike previous work that assumes that devices follow the ideal behavior defined in the specification. We also focus on \emph{transient states} we observe at the RLC level related to promotions and demotions, and their impact on higher-layer performance.   Work by Souders~\cite{souder} measures RRC state machines on client devices, but being a browser-based solution it is not able to account for background activity on the phone or varying network types, measure performance data as accurately as an application can, or observe RLC-level behavior.

Other work has been done on optimizing the use of the cellular network based on 3G or 4G specific phenomena. RadioJockey~\cite{radiojockey} investigates how to effectively trigger fast dormancy based on network traffic patterns, and TailEnder~\cite{tailender} and work by Deng et al.~\cite{trafficaware} propose a method of scheduling data transfers to minimize energy consumption without impacting user-perceived performance.


Related to our RLC fast retransmission proposal, a previous study provides cross comparison analysis over the TCP and RLC protocols, and optimizes the default protocol parameters~\cite{opt.tcp.rlc}. Their primary goal is to improve the performance behavior by introducing TCP congestion control mechanism into the RLC layer. However, their traces are purely generated from simulation software, whereas we use real traffic traces. 

More generally, work has been done to measure performance characteristics of cellular networks and networks on mobile devices. The Livelab project~\cite{livelab} also makes use of users running a measurement app on their phones. A wide range of findings on how users interact with mobile devices have been published, including  measuring web usage in the wild~\cite{livelab-webusage}. Work by Halepovic et al.,~\cite{http_measure} presents a method of passively measuring HTTP transaction latency. Work by Gember et al.,~\cite{incontext} determines how to accurately measure user-perceived performance on user devices. JamLogger~\cite{jamlogger} is an ongoing project to collect general performance and user activity on mobile devices. Our work is complementary to these studies.


%Beyond deployments and testbeds: Experiences with public usage on vehicular WiFi hotspots

%Aditya Akella 	University of Wisconsin
%    Obtaining Representative Measurements of Cellular Network Performance 
%    Explores other factors (network traffic, location etc) that effects measurements
%    Identify that a wide range of factors can affect measurements - we care about relative values
%
%Aruna Balasubramanian 	UW
%     Energy Consumption in Mobile Phones: A Measurement Study and Implications for Network Applications 
%        Tail energy: TailEnder: schedule to avoid tail times

%Joel Sommers 	Colgate University
%    Are Smaller Packets Less Likely to be Lost?:
%    for cellular

% Jamlogger: collects a variety of system statistics periodically in the background; does not have the same challenges as a system looking for specific data, both in measurement cost and in scheduling.  Mobiperf also is not concerned with covering location in the same way and does not have the same challenges.

% LiveLab: Measuring Wireless Networks and Smartphone Users in the Field
% MEasure general real-world smartphone usage and wireless networks. Also uses interrupt-based logging, periodic logging, and collecting when idle.  Hitch-hike onto existing wakeups.  Vary frequency based on context.  More general measurements.  They don't describe power in detail, though - it's a short pape
